Project proposal: Explanations for unexplained effect size heterogeneity

### 2023-06-14 supplement/supplement_A.qmd
correction to supplement names...

# Supplement A: varying reliability SD and the effect of N and k

@@

@@

```{r Figure A3, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/supplement/supplement_A_fig3.png")
```

>>>>

# Supplement A: perfect reliability and Pearson's $r$

Figure A1 shows the bias in heterogeneity estimates in the absence of measurement unreliability in primary studies (i.e., reliability = 1).

@@

_Figure A1._ Bias in heterogeneity estimates in the absence of unreliability in primary studies (i.e., reliability = 1). The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations. Columns indicate the nominal true heterogeneity standard deviation $\tau$. Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).

As can be seen, all three models show a positive bias when heterogeneity is zero, as heterogeneity estimates are truncated at zero. However, we also see a positive bias when heterogeneity is above zero for the Pearson's correlation using the HV method (second row). This is due to a dependancy between effect size and sampling variance for Pearson's $r$ as normally computed. That is, the sampling variance is estimated as $\hat{\sigma}^2_i = \frac{(1-r_i^2)^2}{n_i - 1}$. We do not see this effect for the HS method or HV method using Fisher's $z$. In the HS method $r_i$ has been replaced by $\bar{r}$ and for Fisher's $z$ the sampling variance is approximated as $\hat{\sigma}^2_i = 1 / (n_i -3)$, in both cases removing the dependency between effect size and sampling variance. To demonstrate that this is indeed the case, Figure 2 demonstrates how the bias in Pearson's $r$ disappears if we reshuffle the sampling variances among primary studies so that the dependency between effect size and sampling variance disappears.

@@

_Figure A2._ Removal of the dependency between effect size and sampling variance removes the positive bias in heterogeneity estimates for larger true heterogeneity when using Pearson's $r$. The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 1e3 repetitions. Columns indicate the nominal true heterogeneity standard deviation $\tau$. Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).

### 2023-06-14 supplement/supplement_B.qmd
correction to supplement names...

---
format: pdf
---

@@

# Supplement B: perfect reliability and Pearson's $r$

Figure 1 shows the bias in heterogeneity estimates in the absence of measurement unreliability in primary studies (i.e., reliability = 1).

```{r Figure 1, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/supplement/supplement_B_fig1.png")
```

_Figure 1._ Bias in heterogeneity estimates in the absence of unreliability in primary studies (i.e., reliability = 1). The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations. Columns indicate the nominal true heterogeneity standard deviation $\tau$. Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).

As can be seen, all three models show a positive bias when heterogeneity is zero, as heterogeneity estimates are truncated at zero. However, we also see a positive bias when heterogeneity is above zero for the Pearson's correlation using the HV method (second row). This is due to a dependancy between effect size and sampling variance for Pearson's $r$ as normally computed. That is, the sampling variance is estimated as $\hat{\sigma}^2_i = \frac{(1-r_i^2)^2}{n_i - 1}$. We do not see this effect for the HS method or HV method using Fisher's $z$. In the HS method $r_i$ has been replaced by $\bar{r}$ and for Fisher's $z$ the sampling variance is approximated as $\hat{\sigma}^2_i = 1 / (n_i -3)$, in both cases removing the dependency between effect size and sampling variance. To demonstrate that this is indeed the case, Figure 2 demonstrates how the bias in Pearson's $r$ disappears if we reshuffle the sampling variances among primary studies so that the dependency between effect size and sampling variance disappears.

```{r Figure 2, echo=FALSE, out.width = '100%'}
knitr::include_graphics("../figures/supplement/supplement_B_fig2.png")
```

_Figure 2._ Removal of the dependency between effect size and sampling variance removes the positive bias in heterogeneity estimates for larger true heterogeneity when using Pearson's $r$. The x-axis indicates average effect size and the y-axis estimated heterogeneity in standard deviations averaged across 1e3 repetitions. Columns indicate the nominal true heterogeneity standard deviation $\tau$. Due to truncation in Pearson's $r$ or translation to Fisher's $z$ these values may differ from actual true heterogeneity standard deviation (black solid lines).

>>>>

---

format: pdf

---
# Supplement B: converting heterogeneity between Pearson's r and Fisher's z

@@

The table below shows the results of our procedure to convert between studies variance expressed in Pearson's $r$ to between-studies variance expressed in Fisher's $z$ through the use of $I^2$ as described in the main text.

_Table B1_.
```{r}

source("../code/01_compute_tau_fishers_z.r") #all computations were made here

#Gives us the object 'cond'
cond_table <- cond[,c(1, 3, 2, 4, 6)]

knitr::kable(cond_table, booktabs = TRUE, col.names = c("$\\rho$", "N", "$\\tau_r$", "$I^2$", "$\\tau_z$"), escape = FALSE)
```

_Note._ $\rho$ = average effect size, N = (fixed) within-study sample size, $\tau_r$ = between-studies standard deviation in Pearson's $r$, $\tau_z$ between-studies standard deviation in Fisher's $z$. The $I^2$ statistic is here expressed as a proportion.
